Staged Abstract Interpreters

### 2019-03-26 paper/20_prelim.tex
section 4

\begin{lstlisting}
  def num(i: Int): Ans
  def get(σ: R[Store], ρ: R[Env], x: Ident): R[Value]
  def close(ev: Expr => Ans)(λ: Lam, ρ: R[Env]): R[Value]
  def br0(test: R[Value], thn: => Ans, els: => Ans): Ans
  def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
  def ap_clo(ev: Expr => Ans)(fun: R[Value], arg: R[Value]): Ans
\end{lstlisting}

@@

\begin{lstlisting}
  def ap_clo(ev: EvalFun)(f: Value, arg: Value): Ans = f match {
    case CloV(Lam(x, e), ρ: Env) => for {
      α <- alloc(x)
      _ <- set_store(α → arg)
      rt <- ext_env(ev(e))(x → α)
    } yield rt
  }
\end{lstlisting}

>>>>

\begin{lstlisting}
  def num(i: Int): Ans
  def get(σ: R[Store], ρ: R[Env], x: Ident): R[Value]
  def close(ev: Expr => Ans)(λ: Lam, ρ: R[Env]): R[Value]
  def br0(test: R[Value], thn: => Ans, els: => Ans): Ans
  def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
  def ap_clo(ev: Expr => Ans)(rator: R[Value], rand: R[Value]): Ans
\end{lstlisting}

@@

\begin{lstlisting}
  def ap_clo(ev: EvalFun)(rator: Value, rand: Value): Ans = rator match {
    case CloV(Lam(x, e), ρ: Env) => for {
      α <- alloc(x)
      _ <- set_store(α → rand)
      rt <- ext_env(ev(e))(x → α)
    } yield rt
  }
\end{lstlisting}

### 2019-03-26 paper/30_stageconc.tex
section 4

Readers may notice that the conversion between unstaged monads and monads is 
simply changing the unstaged data to staged data, which in fact can be achieved without
changing the implementation of monads. This is true so far, but as we will see
it is not that easy when introducing the nondeterminism monad (@ListT@) for
abstract interpreters, as the whole list will be a next-stage value instead of
merely the elements in the list.
For this reason, we explicitly distinguish and present the two version of
monads, as a more consistent approach.

\paragraph{Primitive Operations} Most of the primitive operations can be easily
translated to staged version -- we just need to let them works on staged data.
As we mentioned before, we elaborate how to handle functions and applications in
detail. One of the desired goal is to eliminate the monadic layers through
staging, this becomes tricky for clsoures, because we need to actually compile
the lambda term instead of generating a @CloV@ value. To make this happen, when
closing the lambda term with the environment, we need to staging into the body
expression by \textit{collpasing} the @AnsM@ monads to normal values.
The following function @close@ illustrates the idea:
@f@ is a current stage function but takes two next-stage values @v@ and
$\sigma$, inside of which, we eagerly collapse the monads to values by
@ev(e)(ρ_*)(σ_*)@, which has type @Rep[(Value, Store)]@.
Then we generate an intermediate representation of the compiled closure using
@emit_compiled_clo@. The intermediate representation stands for a next-stage
value, and contains the compiled function, which will be generated as a
next-stage function.

>>>>

Readers may notice that the conversion between unstaged monads and staged monads
is simply changing the unstaged data to staged data, which in fact can be
achieved without modifying the implementation of monads. This is true so far, but
as we will see, it is not that easy when introducing the nondeterminism monad
(@ListT@) for abstract interpreters, as the whole list will be a next-stage
value instead of merely the elements in the list. For this reason, we explicitly
distinguish and present the two versions of monads, as a more consistent
approach.

\paragraph{Primitive Operations} Most of the primitive operations can be easily
translated to their staged versions -- we just need to let them works on staged data.
As we mentioned before, we elaborate how to handle functions and applications in
detail. One of the desired goal is to eliminate the monadic layers through
staging, this becomes a little bit tricky for clsoures, because we need actually
compiling the lambda term instead of generating a @CloV@ value, but the @ev(e)@
returns a monad and prevents the specialization. To make this happen, when
closing a lambda term, we need to staging into the body expression by
\textit{collpasing} the @AnsM@ monads to normal values. The
following function @close@ illustrates the idea: @f@ is a current stage function
but takes two next-stage values @v@ and $\sigma$, inside of which, we eagerly
collapse the monads to values by @ev(e)(ρ_*)(σ_*)@, which has type
@Rep[(Value, Store)]@. Then we generate an intermediate representation of the compiled
closure using @emit_compiled_clo@. The intermediate representation stands for a
next-stage value, and contains the compiled function, which will be generated as
a next-stage function.

### 2019-03-26 paper/40_absinterp.tex
section 4

After seeing both unstaged and staged concrete interpreters, now we trun our
focus on abstract interpreters under the same framework. We will first present a
lattice framework with binding-time annotations, but for the purpose of clearly
illustrating the idea, we just use simple abstract domains such as power set
lattice or point-wise map lattice. Then we show the abstract components such as
abstract environments and abstract stores.
The abstract interpreter we present in this section is similar to
\citet{DBLP:journals/pacmpl/DaraisLNH17}'s, it is also
context/path/flow-insensitive by our choice.  In Section \ref{cfa}, we will
instantiate a context-sensitive analysis;
\citet{Darais:2015:GTM:2814270.2814308} shows how to achieve path- and
flow-sensitivity by varying the monads, which is also applicable to our unstaged
or staged abstract interpreter.

@@

A complete lattice over set $S$ is a tuple of $\langle S, \sqsubseteq, \top,
\bot, \sqcup, \sqcap \rangle$, where $\sqsubseteq : S \to S \to Bool$ is the
ordering relation, $\sqcup: S \to S \to S$ is the join (least upper bound)
operator, and $\sqcap: S \to S \to S$ is the meet (greatest lower bound)
operator. The trait @SPLattice@ defines the type class of stage polymorphic
lattices. Similar to @MonadOps@ in Section \ref{generic_if}, we introduce an
additional higher-kinded type @R[_]@ to annotate the binding-times.

\begin{lstlisting}
  trait SPLattice[S, R[_]] {
    val ⊤: R[S];  val ⊥: R[S]
    def ⊑(l1: R[S], l2: R[S]): R[Boolean]
    def ⊔(l1: R[S], l2: R[S]): R[S]
    def ⊓(l1: R[S], l2: R[S]): R[S]
  }
  trait Lattice[S] extends SPLattice[S, NoRep]
\end{lstlisting}

In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
unstaged lattice type class. An simple instance of lattice we used is the power
set abstract domain, where the ordering is subset relation, the join is union
operation, and the meet is intersect operation. The power set lattice can be
implemented as follows:

\begin{lstlisting}
  def SetLattice[T]: Lattice[Set[T]] = new Lattice[Set[T]] {
    lazy val ⊤: Set[T] = Set[T]()
    lazy val ⊥: Set[T] = throw new NotImplementedError("No representation for top")
    def ⊑(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf  l2
    def ⊔(l1: Set[T], l2: Set[T]): Set[T]  = l1 union     l2
    def ⊓(l1: Set[T], l2: Set[T]): Set[T]  = l1 intersect l2
  }
\end{lstlisting}

Other lattices used in the rest of the paper, such as products and maps, can be
implemented similarly, or lifted from existing lattices element-wise or point-wise.

@@

We follow the \textit{abstracting definitional interpreters} (ADIs) approach
\cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete interpreter to
an abstract interpreter. We first widen the concrete values to @AbsValue@: numbers
are lifted to a singleton object @IntTop@ representing the set of all numbers;
closures remain the same. Then the type @Value@ is sets of @AbsValue@ (we abuse
the type names a little bit, as the generic interface defined the abstract type
@Value@).
The address space is constrained to be finited to ensure the computatbility of
analysis -- for the simplest monovariant analysis, we use variable names for
addresses. After setteling the @Value@ and @Addr@ type, the environments and
stores are automatically lifted to abstract environments and abstract stores,
i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@ respectively.
It is importatnt that the abstract store now maps addresses to sets of abstract
values, meaning an address may points to an over-approximated set of values
occurred at runtime.

@@

To prevent the abstract interpreter from falling into nontermination, the ADI
approach uses a two-cache mechanism. Here we first provides the type
definitions, and later we will describe the algorithm in detail. A @Cache@ is a
mapping from configurations @Config@ to sets of value-store pairs, where the
configuration is a triple of current expression being evaluated, environment and
store.

@@

As of now, the store is a mapping from addresses to sets of abstract values.
Therefore, when applying a function, for instance @f(a)@, we may retrieve a set
of closures by dereferencing the address of @f@. We need to apply all of the
target closures, where the nondeterminism is introduced. Another nondeterminism
happens when evaluating conditionals, we have to evaluate both branches if the
condition can not be determined. In other words, we will have multiple computing
results.

@@

\paragraph{Monads for Abstract Interpretation} Now we incorporate the
nondeterminism and caching into the monad stack. For concrete evaluation, we
already have a reader effect for environments and a state effect for stores. We
add three more effects based on it: the nondeterminism effect is represented by
the @ListT[M[_], A]@ monad transformer, inside of which, we have a @ReaderT@ for
a @in@ cache, and a @StateT@ for an @out@ cache.
\todo{highlight where changed}

@@

The sketch implementation of @ListT@ is shown below. The value @run@ it encapsulates
is of type @M[List[A]]@, where @M[_]@ is the inner monad, and @A@ the element
type of the list.

\begin{lstlisting}
  case class ListT[M[_]: Monad, A](run: M[List[A]]) { ... }
\end{lstlisting}

\paragraph{Primitive Operations} \todo{wip}

The operations @get@ and @put@ on abstract stores are slightly changed according
to the abstract semantics: @get@ uses the bottom element in the @Value@ lattice
if the queried address does not exist; @put@ performs the update by joining with
the existing values. function @alloc@ simply invokes the @Addr@ constructor.
\texttt{close} and \texttt{num} lift syntactic literals to our abstract domain,
i.e., a singleton set that contains only one \texttt{CloV} or \texttt{NumV}
object.

\begin{lstlisting}
\end{lstlisting}

For branching, @branch0@ joins the answers from two branches since we are doing
a path-insensitive analysis. For arithmetics, @prim_eval@ returns the top
abstract number element. It is worth noting that @apply_closure@ is where we
handle the non-determinism of applications. The non-determinism happens because
the abstract store maps addresses to sets of abstract values; when dereferencing
an address from the store, we need to explore all possible values to achieve a
sound analysis. In the case of @apply_closure@, the first argument @fs@ may
contain multiple target closures, so we use a @for@ comprehension over all
closures, apply them respectively and form a set result values @vs@; meanwhile,
every time when evaluating the body expression @e@ we use the latest store
$\sigma$@_*@ which is updated iteratively during the loop. Thus this is a
store-widened analysis.

\paragraph{Caching and Fixpoint Iteration} Caching

@@

We described the abstract semantics modularly in the last section, however, the
abstract interpreter may not terminate for some input program. In this section,
we use a memoization technique to ensure its termination. This technique is also
widely used under the name of co-inductive caching algorithm
\cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800} or
truncated depth-first evaluation \cite{Rosendahl:AbsIntPL}.

The idea is to set up two caches called @in@ and @out@, which are both mapping
from the arguments of the interpreter (@Config@) to the result values of the
interpreter (@Ans@). The @in@ cache contains what we already have computed from
the last iteration, the @out@ cache is what we will have computed after this
iteration joined with the previous one. When starting a new iteration, @in@ is
set to be the result of the last iteration, i.e., @out@; @out@ is set to be
empty. In the iteration, we first check whether @out@ contains what we want, if
yes then it is returned; otherwise, we retrieve what we have from @in@, compute
the result for this iteration, and put the joined result back into @out@. Note
that we also instrument the recursive call by putting @cached_ev@ into @evev@.
After one iteration, if @in == out@, then there is no more information to be
gained, thus the iteration should end and we have reached the fixed point.

Finally, we override the definition of @eval_top@ by instantiating @CacheFix@
with @eval@ and starting the first iteration.

>>>>

After seeing both unstaged and staged concrete interpreters, now we trun our
focus on abstract interpreters under the same framework. We will first present a
lattice framework with binding-time abstraction, but for the purpose of clearly
illustrating the idea, we will just use simple abstract domains such as power set
lattice or point-wise map lattice. Then we show the abstract components, namely
abstract values, abstract environments and abstract stores.
The abstract interpreter we present in this section is similar to
\citet{DBLP:journals/pacmpl/DaraisLNH17}'s, it is also
context/path/flow-insensitive by our choice.  In Section \ref{cfa}, we will
instantiate a commonly-used context-sensitive analysis;
\citet{Darais:2015:GTM:2814270.2814308} shows how to achieve path- and
flow-sensitivity by varying the monads, which is also applicable to our unstaged
or staged abstract interpreter.

@@

A complete lattice over set $S$ is a tuple of $\langle S, \sqsubseteq, \top,
\bot, \sqcup, \sqcap \rangle$, where $\sqsubseteq : S \to S \to Bool$ is the
ordering relation, $\sqcup: S \to S \to S$ is the join (least upper bound)
operator, and $\sqcap: S \to S \to S$ is the meet (greatest lower bound)
operator. The trait @SPLattice@ (Figure \ref{fig:splattice}) defines the type
class of stage polymorphic lattices. Similar to @MonadOps@ in Section
\ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to
annotate the binding-times, where @R[_]@ wraps @S@ and @Boolean@ in these
operations.

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
  \begin{lstlisting}[style=small]
trait SPLattice[S, R[_]] {
  val ⊤: R[S];  val ⊥: R[S]
  def ⊑(l1: R[S], l2: R[S]): R[Boolean]
  def ⊔(l1: R[S], l2: R[S]): R[S]
  def ⊓(l1: R[S], l2: R[S]): R[S]
}
trait Lattice[S] extends SPLattice[S, NoRep]
  \end{lstlisting}
  \caption{trait \texttt{SPLattice} and \texttt{Lattice}} \label{fig:splattice}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.6\textwidth}
\begin{lstlisting}[style=small]
def SetLattice[T]: Lattice[Set[T]] = new Lattice[Set[T]] {
  lazy val ⊥: Set[T] = Set[T]()
  lazy val ⊤: Set[T] = error("No representation for ⊤")
  def ⊑(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf  l2
  def ⊔(l1: Set[T], l2: Set[T]): Set[T]  = l1 union     l2
  def ⊓(l1: Set[T], l2: Set[T]): Set[T]  = l1 intersect l2
}
\end{lstlisting}
  \caption{The power set instance for lattice} \label{fig:powerset}
\end{subfigure}
\end{figure}

In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
unstaged lattice type class. An simple instance of lattice we used is the power
set abstract domain (shown in Figure \ref{fig:powerset}), where the ordering is subset relation, the join is union
operation, and the meet is intersect operation.
Other lattices used in the rest of the paper, such as products and maps, can be
implemented similarly, or lifted from existing lattices element-wise or
point-wise.

@@

We follow the \textit{abstracting definitional interpreters} (ADIs) approach
\cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete interpreter to
an abstract interpreter. We first widen the concrete values to @AbsValue@: numbers
are lifted to a singleton object @IntTop@ representing the set of all numbers;
closures remain the same. Then the type @Value@ is sets of @AbsValue@ (we abuse
the type names a little bit, as the generic interface defined the abstract type
@Value@).
The address space is constrained to be finited to ensure the computatbility of
analysis -- for the simplest monovariant analysis, we use variable names for
addresses. After setteling the @Value@ and @Addr@ type, the environments and
stores are automatically lifted to abstract environments and abstract stores,
i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@ respectively.
It is importatnt that the abstract store now maps addresses to sets of abstract
values, meaning an address may point to an over-approximated set of values
occurred at runtime.

@@

To prevent the abstract interpreter from falling into nontermination, the ADI
approach uses a two-cache mechanism. Here we first provide the type
definitions, and later we will describe the algorithm in detail. A @Cache@ is a
mapping from configurations @Config@ to sets of value-store pairs, where the
configuration is a triple of current expression being evaluated, environment and
store.

@@

As of now, the store is a mapping from addresses to sets of abstract values.
Therefore, when applying a function, for instance @f(a)@, we may retrieve a set
of closures by dereferencing the address of @f@. We need to nondeterministically
apply all of the target closures.
Another nondeterminism happens when evaluating conditionals, we have to evaluate
both branches if the condition can not be determined. In other words, we will
have multiple computing results.

@@

\paragraph{Monads for Abstract Interpretation} Now we incorporate the
nondeterminism and caching into the monad stack. For concrete evaluation, we
already have a reader effect for environments and a state effect for stores. We
further add three more effects based on it: the nondeterminism effect is
represented by the @ListT[M[_], A]@ monad transformer, inside of which, we have
a @ReaderT@ for a @in@ cache, and a @StateT@ for an @out@ cache. \todo{highlight
where changed}

@@

The sketch implementation of @ListT@ transformer is shown below. The value @run@
it encapsulates is of type @M[List[A]]@, where @M[_]@ is the inner monad, and
@A@ the element type of the list. The @flatMap@ method first unwraps the inner
monadic value @M[List[A]]@ and obtain a value of @List[A]@, then applies @f@ to
each element in that list, and finally folds the results back to a single
@List[M, B]@.

\begin{lstlisting}
  case class ListT[M[_]: Monad, A](run: M[List[A]]) {
    def flatMap[B](f: A => ListT[M, B]): ListT[M, B] = ...
  }
\end{lstlisting}

\paragraph{Primitive Operations} Some of the primitive operations are changed
according to the new monad stack. A notable change is that the store update
operation joins the value with the existing values, for the same address. Under
abstract interpretation, we may have to explore the results from both branches
of a conditional expression for soundness. In @br0@, we combine the results
using @mplus@ from @MonadPlus@, which requires the values are join-semilattices
(in our case, @List@s and @Map@s).

\begin{lstlisting}
  def set_store(αv: (Addr, Value)): AnsM[Unit] = liftM(StateTMonad.mod(σ => σ ⊔ Map(αv)))
  def br0(test: Value, thn: => Ans, els: => Ans): Ans = ReaderTMonadPlus.mplus(thn, els)
\end{lstlisting}

The value representation of lambda terms is stil @CloV@, but we lift it to a
singleton set.

\begin{lstlisting}
  def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = Set(CloV(λ, ρ))
\end{lstlisting}

For function applications, @ap_clo@ loooks roughly the same as the concrete
interpreter. The difference is now the function @rator@ is a set of closures, so
we can not directly destruct it by pattern matching. Instead, we use an
auxiliary function @lift_nd@ that takes a list and lifts the elements into the
monad stack. Then we can still implement @ap_clo@ in monadic style, where the
@CloV@ comes from the @ListT@ and the nondeterminism is naturally handled.

\begin{lstlisting}
  def lift_nd[T](vs: List[T]): AnsM[T]
  def ap_clo(ev: EvalFun)(rator: Value, rand: Value): Ans = for {
    CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](f.toList)
    α <- alloc(x)
    _ <- set_store(α → rator)
    rt <- ext_env(ev(e))(x → α)
  } yield rt
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph{Caching and Fixpoint Iteration}

As we mentioned ealier, the ADI approach uses a two-cache mechanism to
compute the least fixed-point and prevent nontermination, which is also called a
\textit{co-inductive} caching algorithm or \textit{truncated depth-first
evaluation} \cite{Rosendahl:AbsIntPL}. It has been widely used in abstract
interpretation or fixed point computation
\cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
Rosendahl:AbsIntPL}. Since the @in@ and @out@ cache are wrapped in the monad
stack, we first defined several methods to obtain or update the @in@ and @out@
cache through the monad stack.

@@

Figure \ref{fig:coind_cache} shows the co-inductive caching algorithm that
closes the open-recursive @eval@.
The idea is to first set up two caches called @in@ and @out@ both of type
@Cache@. Then similar to Kleene's iteration algorithm, we compute the analysis
result iteratively. The @in@ cache contains what we already have computed from
the last iteration, and the @out@ cache is what we would have computed after this
iteration joined with the contents from @in@ cache.
When starting a new iteration, @in@ is set to be the result of the last
iteration, i.e., @out@; and @out@ is assigned to be empty.

During the iteration, we first check whether @out@ contains the configuration
@cfg@, representing the current desired computation, if
yes then result is directly returned through the monads.

Otherwise, we first retrieve the results from @in@ (@⊥@ if not contained in
@in@), and put it into the @out@ cache.
Then we invoke @ev@ to compute the result for this iteration.
After the evaluation, the store @σ@ may have been changed. So we update the
@out@ cache with the new result, i.e., @(v, σ)@. Similar to store update, cache
update is also point-wise join.
Note that for the recursive call of @ev@, we also instrument it by puting
@fix(ev)@ as the first argument to @ev@.

The iteration terminates when the resulting @out@ cache is equivalent to the
input @in@ cache, which means there is no more analysis result can be gained,
thus the iteration should end and we have reached the least fixed point.

\begin{figure}[t!]
  \centering
\begin{lstlisting}
  def fix(ev: EvalFun => EvalFun): EvalFun = e => for {
    ρ <- ask_env; σ <- get_store; in <- ask_in_cache; out <- get_out_cache; val cfg = (e, ρ, σ)
    rt <- if (out.contains(cfg)) for {
            (v, s) <- lift_nd[(Value, Store)](out(cfg).toList)
             _ <- put_store(s)
          } yield v
          else for {
            _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
            v <- ev(fix(ev))(e)
            σ <- get_store
            _ <- update_out_cache(cfg, (v, σ))
          } yield v
  } yield rt
\end{lstlisting}
\vspace{-1em}
\caption{The unstaged co-inductive caching algorithm.}
\label{fig:coind_cache}
\end{figure}

\begin{lstlisting}
  def run(e: Expr): Result = fix(eval)(e)(ρ0)(σ0)(cache0)(cache0).run
\end{lstlisting}

Finally, we override the top-level @run@ method by running the monad with
the empty abstract environment $\rho_0$, initial store $\sigma_0$, and initial
caches @cache@$_0$.

### 2019-03-26 paper/60_cases.tex
section 4

\begin{figure*}
  \centering
  \begin{subfigure}[h]{0.49\textwidth}
    \centering
    \begin{lstlisting}[
      language=Scala,%
      mathescape=true,%
      %  columns=[c]fixed,%
      %  basewidth={0.5em, 0.40em},%
      aboveskip=2pt,%\smallskipamount,
      belowskip=1pt,%\negsmallskipamount,
      lineskip=-1pt,
      basewidth={0.6em, 0.45em},%
      %  backgroundcolor=\color{listingbg},
      basicstyle=\fontsize{6}{8}\selectfont\ttfamily,
      keywordstyle=\keywordstyle,
      commentstyle=\commentstyle,
      stringstyle=\stringstyle,
    ]
type CompAnalysis = Store => Store
def compProgram(prog: Expr): CompAnalysis = compCall(prog)
def compCall(call: Expr): CompAnalysis = call match {
  case Letrec(bds, body) =>
    val C1 = compCall(body); val C2 = compArgs(bds.map(_.value))
    ($\sigma$: Store) => C1(C2($\sigma$.update(bds.map(_.name), 
       bds.map(b => Set(b.value.asInstanceOf[Lam])))))
  case App(f, args) =>
    val C1 = compApp(f, args); val C2 = compArgs(args)
    ($\sigma$: Store) => C1(C2($\sigma$))
}
def compApp(f: Expr, args: List[Expr]): CompAnalysis = 
  f match {
    case Var(x) => ($\sigma$: Store) => 
      analysisAbsApp($\sigma$.lookup(x), args, $\sigma$)
    case Op(_) => compArgs(args)
    case Lam(vars, body) =>
      val C = compCall(body)
      ($\sigma$: Store) => C($\sigma$.update(vars, args.map(primEval(_, $\sigma$))))
  }
def compArgs(args: List[Expr]): CompAnalysis = args match {
  case Nil => ($\sigma$: Store) => $\sigma$
  case (arg@Lam(vars, body))::rest =>
    val C1 = compCall(body); val C2 = compArgs(rest)
    ($\sigma$: Store) => C2(C1($\sigma$))
  case _::rest => compArgs(rest)
}
  \end{lstlisting}
  \end{subfigure}
\hfill
  \begin{subfigure}[h]{0.49\textwidth}
    \centering
    \begin{lstlisting}[
      language=Scala,%
      mathescape=true,%
      %  columns=[c]fixed,%
      %  basewidth={0.5em, 0.40em},%
      aboveskip=2pt,%\smallskipamount,
      belowskip=1pt,%\negsmallskipamount,
      lineskip=-1pt,
      basewidth={0.6em, 0.45em},%
      %  backgroundcolor=\color{listingbg},
      basicstyle=\fontsize{6}{8}\selectfont\ttfamily,
      keywordstyle=\keywordstyle,
      commentstyle=\commentstyle,
      stringstyle=\stringstyle,
    ]
def analyzeProgram(prog: Expr, $\sigma$: Rep[Store]): Rep[Store] = 
  analyzeCall(prog, $\sigma$)
def analyzeCall(call: Expr, $\sigma$: Rep[Store]): Rep[Store] = 
  call match {
    case App(f, args) => analyzeApp(f, args, analyzeArgs(args, $\sigma$))
    case Letrec(bds, body) =>
      val $\sigma$_* = $\sigma$.update(bds.map(_.name), 
        bds.map(b => Set(b.value.asInstanceOf[Lam])))
      val $\sigma$_** = analyzeArgs(bds.map(_.value), $\sigma$_*)
      analyzeCall(body, $\sigma$_**)
  }
def analyzeApp(f: Expr, args: List[Expr], $\sigma$: Rep[Store]): 
  Rep[Store] = f match {
    case Var(x) => analyzeAbsApp(args, $\sigma$(x), $\sigma$)
    case Op(_) => analyzeArgs(args, $\sigma$)
    case Lam(vars, body) =>
      val $\sigma$_* = $\sigma$.update(vars, args.map(primEval(_, $\sigma$)))
      analyzeCall(body, $\sigma$_*)
  }
def analyzeArgs(args: List[Expr], $\sigma$: Rep[Store]): Rep[Store] = 
  args match {
    case Nil => $\sigma$
    case Lam(vars, body)::rest => 
      analyzeArgs(rest, analyzeCall(body, $\sigma$))
    case _::rest => analyzeArgs(rest, $\sigma$)
  }
  \end{lstlisting}

>>>>

\begin{figure*}
  \centering
  \begin{subfigure}[h]{0.49\textwidth}
    \centering
    \begin{lstlisting}[style=extrasmall]
type CompAnalysis = Store => Store
def compProgram(prog: Expr): CompAnalysis = compCall(prog)
def compCall(call: Expr): CompAnalysis = call match {
  case Letrec(bds, body) =>
    val C1 = compCall(body); val C2 = compArgs(bds.map(_.value))
    ($\sigma$: Store) => C1(C2($\sigma$.update(bds.map(_.name), 
       bds.map(b => Set(b.value.asInstanceOf[Lam])))))
  case App(f, args) =>
    val C1 = compApp(f, args); val C2 = compArgs(args)
    ($\sigma$: Store) => C1(C2($\sigma$))
}
def compApp(f: Expr, args: List[Expr]): CompAnalysis = 
  f match {
    case Var(x) => ($\sigma$: Store) => 
      analysisAbsApp($\sigma$.lookup(x), args, $\sigma$)
    case Op(_) => compArgs(args)
    case Lam(vars, body) =>
      val C = compCall(body)
      ($\sigma$: Store) => C($\sigma$.update(vars, args.map(primEval(_, $\sigma$))))
  }
def compArgs(args: List[Expr]): CompAnalysis = args match {
  case Nil => ($\sigma$: Store) => $\sigma$
  case (arg@Lam(vars, body))::rest =>
    val C1 = compCall(body); val C2 = compArgs(rest)
    ($\sigma$: Store) => C2(C1($\sigma$))
  case _::rest => compArgs(rest)
}
  \end{lstlisting}
  \end{subfigure}
\hfill
  \begin{subfigure}[h]{0.49\textwidth}
    \centering
    \begin{lstlisting}[style=extrasmall]
def analyzeProgram(prog: Expr, $\sigma$: Rep[Store]): Rep[Store] = 
  analyzeCall(prog, $\sigma$)
def analyzeCall(call: Expr, $\sigma$: Rep[Store]): Rep[Store] = 
  call match {
    case App(f, args) => analyzeApp(f, args, analyzeArgs(args, $\sigma$))
    case Letrec(bds, body) =>
      val $\sigma$_* = $\sigma$.update(bds.map(_.name), 
        bds.map(b => Set(b.value.asInstanceOf[Lam])))
      val $\sigma$_** = analyzeArgs(bds.map(_.value), $\sigma$_*)
      analyzeCall(body, $\sigma$_**)
  }
def analyzeApp(f: Expr, args: List[Expr], $\sigma$: Rep[Store]): 
  Rep[Store] = f match {
    case Var(x) => analyzeAbsApp(args, $\sigma$(x), $\sigma$)
    case Op(_) => analyzeArgs(args, $\sigma$)
    case Lam(vars, body) =>
      val $\sigma$_* = $\sigma$.update(vars, args.map(primEval(_, $\sigma$)))
      analyzeCall(body, $\sigma$_*)
  }
def analyzeArgs(args: List[Expr], $\sigma$: Rep[Store]): Rep[Store] = 
  args match {
    case Nil => $\sigma$
    case Lam(vars, body)::rest => 
      analyzeArgs(rest, analyzeCall(body, $\sigma$))
    case _::rest => analyzeArgs(rest, $\sigma$)
  }
  \end{lstlisting}
